---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

library(tidyverse)
library(hrbrthemes)
library(echor)
library(extrafont)
loadfonts()
```

# echor

[![Travis build status](https://travis-ci.org/mps9506/echor.svg?branch=master)](https://travis-ci.org/mps9506/echor)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/mps9506/echor?branch=master&svg=true)](https://ci.appveyor.com/project/mps9506/echor)
[![Coverage status](https://codecov.io/gh/mps9506/echor/branch/master/graph/badge.svg)](https://codecov.io/github/mps9506/echor?branch=master)

## Overview

echor downloads discharge and emission data for EPA permitted facilities using the [EPA ECHO API](https://echo.epa.gov/).

## Installation

```{r Installation, eval=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

devtools::install_github("mps9506/echor")

```

## Usage

Coming soon.

## Examples

### Download information about facilities with an NPDES permit

We can look up plants by permit id, bounding box, and numerous other parameters. I plan on providing documentation of available parameters. However, arguments can be looked up here: [get_cwa_rest_services_get_facility_info](https://echo.epa.gov/tools/web-services/facility-search-water#!/Facility_Information/get_cwa_rest_services_get_facility_info)

```{r example1, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(echor)

## echo*GetFacilityInfo() will return a dataframe or simple features (sf) dataframe.

df <- echoWaterGetFacilityInfo(output = "df", 
                               xmin = '-96.387509', 
                               ymin = '30.583572', 
                               xmax = '-96.281422', 
                               ymax = '30.640008')

head(df)
```

When returned as sf dataframes, the data is suitable for immediate spatial plotting or analysis:
```{r example2, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggmap)
library(sf)
library(ggrepel)
## This example requires the development version of ggplot with support
## for geom_sf()
## and uses theme_ipsum_rc() from library(hrbrthemes)


df <- echoWaterGetFacilityInfo(output = "sf", 
                               xmin = '-96.387509', 
                               ymin = '30.583572', 
                               xmax = '-96.281422', 
                               ymax = '30.640008')

collegestation <- get_map(location = c(-96.387509, 30.583572,
                                       -96.281422, 30.640008), 
                          zoom = 14, maptype = "toner")

##to make labels, need to map the coords and use geom_text :(
## can't help but think there is an easier way to do this

df <- df %>%
  mutate(
    coords = map(geometry, st_coordinates),
    coords_x = map_dbl(coords, 1),
    coords_y = map_dbl(coords, 2)
  )

ggmap(collegestation) + 
  geom_sf(data = df, inherit.aes = FALSE, shape = 21, 
          color = "darkred", fill = "darkred", 
          size = 2, alpha = 0.25) +
  geom_label_repel(data = df, aes(x = coords_x, y = coords_y, label = SourceID),
                   point.padding = .5, min.segment.length = 0.1,
                   size = 2, color = "dodgerblue") +
  theme_ipsum_rc() +
  labs(x = "Longitude", y = "Latitude", 
       title = "NPDES permits near Texas A&M",
       caption = "Source: EPA ECHO database")

```

### Download discharge/emissions data

Use `echoGetEffluent()` or `echoGetCAAPR()` to download tidy dataframes of permitted water discharger Discharge Monitoring Report (DMR) or permitted emitters Clean Air Act annual emissions reports.

```{r}
df <- echoGetEffluent(p_id = 'tx0119407', parameter_code = '00300')

df <- df %>%
  filter(!is.na(DMRValueNmbr) & ValueTypeCode == "C1")

ggplot(df) +
  geom_line(aes(MonitoringPeriodEndDate, DMRValueNmbr)) +
  theme_ipsum_rc(grid = "Y") +
  labs(x = "Monitoring period date",
       y = "Dissolved oxygen concentration (mg/l)",
       title = "Reported minimum dissolved oxygen concentration",
       subtitle = "NPDES ID = TX119407",
       caption = "Source: EPA ECHO")

```

