---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.retina = 2
)

library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(echor)
library(extrafont)
loadfonts()
```

# echor
<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/echor)](https://cran.r-project.org/package=echor)

[![R build status](https://github.com/mps9506/echor/workflows/R-CMD-check/badge.svg)](https://github.com/mps9506/echor/actions)
[![Travis build status](https://travis-ci.org/mps9506/echor.svg?branch=master)](https://travis-ci.org/mps9506/echor)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/mps9506/echor?branch=master&svg=true)](https://ci.appveyor.com/project/mps9506/echor)
[![Coverage status](https://codecov.io/gh/mps9506/echor/branch/master/graph/badge.svg)](https://codecov.io/github/mps9506/echor?branch=master)
[![DOI](https://zenodo.org/badge/122131508.svg)](https://zenodo.org/badge/latestdoi/122131508)
<!-- badges: end -->

## Overview

echor downloads wastewater discharge and air emission data for EPA permitted facilities using the [EPA ECHO API](https://echo.epa.gov/).

## Installation

echor is on CRAN:

```{r Install, eval=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
install.packages("echor")
```

Or install the development version from github:

```{r InstallDev, eval=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
remotes::install_github("mps9506/echor")
```

## Usage

[Getting started](https://mps9506.github.io/echor/articles/introduction.html)

[Functions](https://mps9506.github.io/echor/reference/index.html)

## Examples

### Download information about facilities with an NPDES permit

We can look up plants by permit id, bounding box, and numerous other parameters. I plan on providing documentation of available parameters. However, arguments can be looked up here: [get_cwa_rest_services_get_facility_info](https://echo.epa.gov/tools/web-services/facility-search-water#!/Facility_Information/get_cwa_rest_services_get_facility_info)

```{r example1, echo=TRUE, message=FALSE, warning=FALSE}
library(echor)

## echoWaterGetFacilityInfo() will return a dataframe or simple features (sf) dataframe.

df <- echoWaterGetFacilityInfo(output = "df", 
                               xmin = '-96.387509', 
                               ymin = '30.583572', 
                               xmax = '-96.281422', 
                               ymax = '30.640008',
                               p_ptype = "NPD")

head(df)
```

The ECHO database can provide over 270 different columns. echor returns a subset of these columns that should work for most users. However, you can specify what data you want returned. Use `echoWaterGetMeta()` to return a dataframe with column numbers, names, and descriptions to identify the columns you want returned. Then include the column numbers as a comma separated string in the `qcolumns` argument. In the example below, the `qcolumns` argument indicates the dataframe will include plant name, 8-digit HUC, latitude, longitude, and total design flow.

```{r example2, echo=TRUE, message=FALSE, warning=FALSE}
df <- echoWaterGetFacilityInfo(output = "df", 
                               xmin = '-96.387509', 
                               ymin = '30.583572', 
                               xmax = '-96.281422', 
                               ymax = '30.640008',
                               qcolumns = '1,14,23,24,25',
                               p_ptype = "NPD")
head(df)
```



When returned as sf dataframes, the data is suitable for immediate spatial plotting or analysis:
```{r example3, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%"}
library(ggspatial)
library(sf)
library(ggrepel)
library(purrr)
## This example requires the development version of ggplot with support
## for geom_sf()
## and uses theme_ipsum_rc() from library(hrbrthemes)


df <- echoWaterGetFacilityInfo(output = "sf", 
                               xmin = '-96.387509', 
                               ymin = '30.583572', 
                               xmax = '-96.281422', 
                               ymax = '30.640008',
                               p_ptype = "NPD")

##to make labels, need to map the coords and use geom_text :(
## can't help but think there is an easier way to do this

df <- df %>%
  mutate(
    coords = map(geometry, st_coordinates),
    coords_x = map_dbl(coords, 1),
    coords_y = map_dbl(coords, 2)
  )

ggplot(df) +
  annotation_map_tile(zoomin = -1, progress = "none") +
  geom_sf(inherit.aes = FALSE, shape = 21, 
          color = "darkred", fill = "darkred", 
          size = 2, alpha = 0.25) +
  geom_label_repel(data = df, aes(x = coords_x, y = coords_y, label = SourceID),
                   point.padding = .5, min.segment.length = 0.1,
                   size = 2, color = "dodgerblue") +
  theme_ipsum(plot_margin = margin(5, 5, 5, 5)) +
  labs(x = "Longitude", y = "Latitude", 
       title = "NPDES permits near Texas A&M",
       caption = "Source: EPA ECHO database")

```

### Download discharge/emissions data

Use `echoGetEffluent()` or `echoGetCAAPR()` to download tidy dataframes of permitted water discharger Discharge Monitoring Report (DMR) or permitted emitters Clean Air Act annual emissions reports. Please note that all variables are returned as *character* vectors.

```{r message=TRUE, warning=FALSE}
df <- echoGetEffluent(p_id = 'tx0119407', parameter_code = '00300')

df <- df %>%
  mutate(dmr_value_nmbr = as.numeric(dmr_value_nmbr),
         monitoring_period_end_date = as.Date(monitoring_period_end_date,
                                              "%m/%d/%Y")) %>%
  filter(!is.na(dmr_value_nmbr) & limit_value_type_code == "C1")

ggplot(df) +
  geom_line(aes(monitoring_period_end_date, dmr_value_nmbr)) +
  theme_ipsum(grid = "Y") +
  labs(x = "Monitoring period date",
       y = "Dissolved oxygen concentration (mg/l)",
       title = "Reported minimum dissolved oxygen concentration",
       subtitle = "NPDES ID = TX119407",
       caption = "Source: EPA ECHO")

```

## Session Info

```{r message=FALSE, warning=FALSE, error=FALSE}
sessioninfo::platform_info()
sessioninfo::package_info()
```

